---
title: "multiWGCNA: visualizing condition-specific networks"
author: "Dario Tommasini"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Astrocyte multiWGCNA network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In this vignette, we will be showing how users can draw a multiWGCNA network (Figure 2C from Tommasini and Fogel. BMC Bioinformatics. 2023.), which highlights the presence or absence of a biological network in the different conditions of their experiment. We will be using the astrocyte Ribotag data from Itoh et al. PNAS. 2018 (https://doi.org/10.1073/pnas.1716032115). This data comes with the multiWGCNA package and can be accessed as shown below. 

# Load multiWGCNA library

```{r setup}
library(multiWGCNA)
```

# Pre-processing

```{r}
# Download data from the ExperimentHub
library(ExperimentHub)
eh = ExperimentHub()

# Note: this requires the SummarizedExperiment package to be installed
eh_query = query(eh, c("multiWGCNAdata"))
astrocyte_se = eh_query[["EH8223"]]

# Collect the metadata in the sampleTable
sampleTable = colData(astrocyte_se)

# Check the data
assays(astrocyte_se)[[1]][1:5, 1:5]
sampleTable

# Define our conditions for trait 1 (disease) and 2 (brain region)
conditions1 = unique(sampleTable[,2])
conditions2 = unique(sampleTable[,3])
```

# Network construction

We now perform network construction, module eigengene calculation, module-trait correlation. Let's use power = 12 since we used this in our paper (Tommasini and Fogel. BMC Bioinformatics. 2023.) for all the networks. 

```{r, eval = FALSE}
# Construct the combined networks and all the sub-networks (EAE, WT, and each region)
# Same parameters as Tommasini and Fogel. BMC Bioinformatics
astrocyte_networks = constructNetworks(astrocyte_se, sampleTable, conditions1, conditions2, 
                                  networkType = "signed", TOMType = "unsigned", 
                                  power = 12, minModuleSize = 100, maxBlockSize = 25000,
                                  reassignThreshold = 0, minKMEtoStay = 0, mergeCutHeight = 0,
                                  numericLabels = TRUE, pamRespectsDendro = FALSE, 
                                  deepSplit = 4, verbose = 3)

```

This step takes a while since it performs seven network constructions, so we also provide the WGCNA object list (astrocyte_networks) in a loadable format. These were generated from the function above. 

```{r}
# Load pre-computed astrocyte networks
astrocyte_networks = eh_query[["EH8222"]] 

# Check one of the WGCNA objects
astrocyte_networks[["combined"]]
```

# The multiWGCNA network

## Compare modules by overlap 

Next, we compare modules (by hypergeometric overlap) across conditions. We'll save the results in a list. 

```{r, fig.height = 5, fig.width = 8}
# Save results to a list
results = list()
results$overlaps = iterate(astrocyte_networks, overlapComparisons, plot=FALSE)

# Check the overlaps, ie between the EAE and wildtype networks
head(results$overlaps$EAE_vs_WT$overlap)
```

## Identify a module of interest

Then, we perform differential module expression analysis to detect modules with disease-associated expression patterns. This incorporates the linear model described in the paper and tests for significance using ANOVA. 

```{r, fig.height = 6, fig.width = 7}
# Run differential module expression analysis (DME) on combined networks
results$diffModExp = runDME(astrocyte_networks[["combined"]], 
                            sampleTable,
                            p.adjust = "fdr", 
                            refCondition = "Region", 
                            testCondition = "Disease") 
                            # plot=TRUE, 
                            # out="ANOVA_DME.pdf")

# Check results sorted by disease association FDR
results$diffModExp[order(results$diffModExp$Disease),]

# You can check the expression of module M13 from Tommasini and Fogel. BMC Bioinformatics. 2023 like this. Note that the values reported in the bottom panel title are p-values and not adjusted for multiple comparisons like in results$diffModExp
diffModuleExpression(astrocyte_networks[["combined"]], 
                     geneList = topNGenes(astrocyte_networks[[1]], "combined_013"), 
                     design = sampleTable,
                     test = "ANOVA",
                     plotTitle = "combined_013",
                     plot = TRUE)

```

## Draw the multiWGCNA network

We can now check to see if M13 is present in any of the sub-networks. An easy way to do this is using the network-network correspondences from hypergeometric overlap. These are stored in results$overlaps. We can plot these in a convenient visualization scheme that also organizes the three levels of the multiWGCNA analysis: 1) combined network, 2) EAE and wildtype networks, and 3) the four regional networks. 

```{r, fig.height = 6, fig.width = 7}
drawMultiWGCNAnetwork(astrocyte_networks, 
                      results$overlaps, 
                      "combined_013", 
                      design = sampleTable, 
                      overlapCutoff = 0, 
                      padjCutoff = 1, 
                      removeOutliers = TRUE, 
                      alpha = 1e-50, 
                      layout = NULL, 
                      hjust = 0.4, 
                      vjust = 0.3, 
                      width = 0.5)

```

We see that M13 is really only present in the EAE network, but not any of the other sub-networks. Most importantly, it cannot be resolved in the wildtype network. This makes M13 a biologically interesting network, both in terms of differential expression and differential co-expression. 

## Follow up with a preservation analysis

Typically, you would want to follow this up with a preservation analysis between EAE and WT (described in general_workflow.Rmd). 

```{r, eval = FALSE, fig.height = 3, fig.width = 7}
# To enable multi-threading
library(doParallel)
library(WGCNA)
nCores = 8
registerDoParallel(cores = nCores)
enableWGCNAThreads(nThreads = nCores)

# Calculate preservation statistics
results$preservation=iterate(astrocyte_networks[conditions1], 
                             preservationComparisons, 
                             write=FALSE, 
                             plot=TRUE, 
                             nPermutations=100)
```

# Conclusion

Then, one can perform a permutation procedure that estimates the probability of observing a disease/wildtype module with this preservation score in the wildtype/disease setting (diseasePreservationPtest). Please see Tommasini and Fogel. BMC Bioinformatics. 2023. for more details. 

```{r}
sessionInfo()
```
